---
title: "CHL5233: Quarto in-class assignment"
author: "Tallinn Splinter"
format: 
  pdf:
    geometry:
      - top=30mm
      - left=20mm
      - heightrounded
---



```{r}
#| echo: false
#| output: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
```


## In Class Assignment - September 28th

For this exercise, we will use the following two data sets.

1. `countries.txt`: A data set that contains a list of countries used in the paper "Implications of armed conflict for maternal and child health: A regression analysis of data from 181 countries for 2000–2019" and their three-digit ISO (International Organization for Standardization) code. 

2. `regions.txt`: A data set that contains a list of all major countries in the world and their
    a. ISO code (`country_code`)
    b. region (`region`)
    c. sub-region (`sub_region`)
    d. numerical region code (`region_code`)
    e. numerical sub-region code (`sub_region_code`)

The **final goal** is to display a **bar chart** that depicts how many countries are in each region. 


## Read in data and create an analytical data set

**First**, edit the code chunk below that reads in `countries.txt` to also display the number of rows and columns of the data. 

```{r}
#| echo: TRUE
countries <- read.table("../data/raw/countries.txt", header = TRUE)
nrow(countries)
ncol(countries)
```
*There are 186 countries in this countries dataset*

**Second**, read in `regions.txt` and also write a sentence that states how many countries are included in the data set.

```{r}
#| echo: TRUE
regions <- read.table("../data/raw/regions.txt", header = TRUE)
nrow(regions)
ncol(regions)
```
*There are 249 countries in this `regions` data set*

**Third**, subset the `regions` data set to only include countries from the `countries` data set.

```{r}
#| echo: TRUE
regions_subset <- regions |>
  filter(country_code %in% countries$ISO)

```



## Tabulate countries by region and sub-region

Next, we want to find out how many countries are in a region. Use the `group_by()` and `summarise()` functions in the `dplyr` package to count the number of countries in each region. Then use the `kable()` function in the `knitr` package to create a simple table with the column headers, "Region" and "Count". 

```{r}
#| echo: TRUE
regions_count <- regions_subset |>
  group_by(region) |>
  summarise(Count = n())


regions_count |>
  kable(col.names = c("Region", "Count"))
```


Create a similar table that shows the number of countries in each **sub-region**.

```{r}
#| echo: TRUE
sub_regions_count <- regions_subset |>
  group_by(sub_region) |>
  summarise(Count = n())

sub_regions_count |>
  kable(col.names = c("Sub-Region", "Count"))
```



## Data visualization using ggplot

Here, we will use `geom_bar()` function within `ggplot()` to create a bar chart that displays the **number of countries** in each **sub-region**. 

1. Color code each sub-region by **region**.
2. Create bars that run horizontally where sub-region names appear along the y-axis. 
3. Order the bars such that the sub-region with the most number countries appear at the top and the sub-region with the least number of countries appear at the bottom. 

```{r}
#| echo: TRUE
ggplot(regions_subset,
       aes(x = reorder(sub_region, sub_region, FUN = length),
           fill = region)) +
  geom_bar() +
  coord_flip() +
  labs(
    x = "Sub-Region",
    y = "Number of Countries",
    fill = "Region"
  )
```

